---
title: "Reproducible Research Project 1"
author: "nhanwei"
date: "15 October 2014"
output: html_document
---

##Loading and preprocessing the data  

First we load our dataset into R which is located in our working directory:  

```{r}
library(ggplot2)
library(gridExtra)
setwd("~/Downloads")
activity <- read.csv("activity.csv", header=T)   # Loading into R 
activity$date <- as.Date(activity$date)         # Change to correct date format
```

##What is mean total number of steps taken per day?  

```{r}
rmNA <- activity[(!is.na(activity$steps)),]     # removed rows with NA 
StepSum <- tapply(rmNA$steps, rmNA$date, sum)   
qplot(StepSum, xlab = "Sum of Steps Taken Each Day", ylab = "Number of Days", 
      main = "Number of Days Against Sum of Steps Taken Each Day", 
      binwidth = 2000)
cat("Mean total number of steps taken per day:", mean(StepSum))   
cat("Median total number of steps taken per day:", median(StepSum)) 
```

##What is the average daily activity pattern?

```{r}
# Calculating Step Averages for each Interval
StepIntAvg <- as.vector(tapply(rmNA$steps, rmNA$interval, mean))
# Taking interval names
IntervalNames <- rmNA$interval[1:288]
# Combining both vectors
timeplot <- data.frame(Interval = IntervalNames, Averages = StepIntAvg)
# Graph Plot
p <- ggplot(timeplot, aes(x=Interval, y=Averages))
p + geom_line() + xlab("Intervals") + 
    ggtitle("Interval Averages Against Interval")
cat("The 5-minute interval that contains the maximum average number of steps:",
    timeplot[timeplot$Averages == max(timeplot$Averages),1])
```

##Imputing missing values

```{r}
cat("Total number of rows with NAs:", sum(is.na(activity$steps)))
# Taking out rows without NAs
NArows <- activity[(is.na(activity$steps)),]
# Replacing NAs with Calculated Averages for the Steps in each Interval
NArows$steps <- StepIntAvg
# Now we combine all rows together
newactivity <- rbind(NArows, rmNA)
# Calculating the sum of steps in each day
NewStepSum <- tapply(newactivity$steps, newactivity$date, sum)  
# Graph Plot
qplot(NewStepSum, xlab = "New Sum of Steps Taken Each Day", 
      ylab = "Number of Days", 
      main = "Number of Days Against New Sum of Steps Taken Each Day", 
      binwidth = 2000)
cat("New Mean total number of steps taken per day:", mean(NewStepSum))   
cat("New Median total number of steps taken per day:", median(NewStepSum)) 
```

The mean and median differs by very little.  
The impact of imputing missing data is that there are more days in the data now.
Previously those 8 days were removed because of the NA rows. Due to the imputing,
these 8 days have the same total daily number of steps which is 10766

##Are there differences in activity patterns between weekdays and weekends?

```{r}
# Creating new factor variable
diffactivity <- data.frame(newactivity, week = ifelse(
    weekdays(newactivity$date) == "Sunday" | weekdays(newactivity$date) == "Saturday"
    , "weekend", "weekday"))
# Separating rows by week variable
WeekendData <- diffactivity[which(diffactivity$week=="weekend"),]
WeekdayData <- diffactivity[which(diffactivity$week=="weekday"),]
# Calculating Step Averages for each Interval
WeekendStepAvg <- as.vector(tapply(WeekendData$steps, WeekendData$interval, 
                                   mean))
WeekdayStepAvg <- as.vector(tapply(WeekdayData$steps, WeekdayData$interval, 
                                   mean))

WeekendTimePlot <- data.frame(Interval = IntervalNames,Averages= WeekendStepAvg)
WeekdayTimePlot <- data.frame(Interval = IntervalNames,Averages= WeekdayStepAvg)

# Graph Plot
plot1 <- qplot(WeekendTimePlot$Interval, WeekendTimePlot$Averages,
               xlab ="Interval", ylab="Averages", geom = "line",
               main ="Weekend")
plot2 <- qplot(WeekdayTimePlot$Interval, WeekdayTimePlot$Averages,
               xlab ="Interval", ylab="Averages",geom = "line",
               main ="Weekday")
grid.arrange(plot1, plot2, ncol=2)
```
